Estimating Fixed Frame Galaxy Magnitudes in the SDSS$^1$

### 2002-07-20 docs/paper/paper.tex
changed major error in overall zeropoint of K-corrections
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@114 75dff457-8751-0410-ab1a-cd701654ce6b

\newcommand{\vv}[1]{{\bf #1}}
\newcommand{\df}{\delta}
\newcommand{\dfft}{{\tilde{\delta}}}
\newcommand{\betaft}{{\tilde{\beta}}}
\newcommand{\erf}{{\mathrm{erf}}}
\newcommand{\erfc}{{\mathrm{erfc}}}
\newcommand{\Step}{{\mathrm{Step}}}
\newcommand{\ee}[1]{\times 10^{#1}}
\newcommand{\avg}[1]{{\langle{#1}\rangle}}
\newcommand{\Avg}[1]{{\left\langle{#1}\right\rangle}}
\def\simless{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'074$}}\mathchar"7218\,$}}} % < or of order
\def\simgreat{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'076$}}\mathchar"7218\,$}}} % > or of order
\newcommand{\iras}{{\sl IRAS\/}}
\newcommand{\petroratio}{{{\mathcal{R}}_P}}
\newcommand{\petroradius}{{{r}_P}}
\newcommand{\petronumber}{{{N}_P}}
\newcommand{\petroratiolim}{{{\mathcal{R}}_{P,\mathrm{lim}}}}
\newcommand{\band}[2]{\ensuremath{^{#1}\!{#2}}}

@@

In the past, people have accounted for these effects using
``$K$-corrections'' applied to the observed magnitudes. In these
analyses, a function $K(z)$ is added to the standard cosmological
bolometric distance modulus $\mathrm{DM}(z)$ to obtain the
relationship between the apparent magnitude $m$ and the absolute
magnitude $M$: 
\begin{equation}
m = M + \mathrm{DM}(z) + K(z)
\end{equation}
Sometimes a single function has been applied regardless of the galaxy
type, though more recently it has become standard to use a discrete
set of $K(z)$ functions depending on galaxy type (based either on
morphology or spectral features). As described in \cite{oke68a} and
later papers, the function is based on the projection of an assumed
galaxy SED redshifted to $z$ onto the measured instrument response as
a function of wavelength (including the effects of the atmosphere, the
reflectivity of the mirrors, the filter transmission, and the response
of the CCD device or photographic emulsion). For example, for the
Sloan Digital Sky Survey (SDSS) filter system, \citet{frei94a} and
\citet{fukugita95a} have both presented results for the
$K$-corrections of galaxies. Each of these papers tabulates $K(z)$
(called $k(z)$ by \citealt{frei94a}) for galaxies of various
morphological types, based on galaxy spectrophotometry.\footnote{Note
that the description on p. 57 of \citet{binney98a} of the meaning of
$k(z)$ in \citet{frei94a} is incorrect; in fact, their $k(z)$ is
exactly equivalent to our $K(z)$.}

@@

We here define a notation which we will use in subsequent papers for
the effective rest-frame bandpass at any redshift in band $b$, namely:
\begin{equation}
\band{z}{b}
\end{equation}
read, ``$b$-band at redshift $z$.'' For example, if we measure the
flux of a galaxy at $z=0.22$ using a $g$-band filter, we have obtained
the $\band{0.22}{g}$-band magnitude of the galaxy. The goal of
``$K$-correction'' is to convert the $ugriz$ bands at redshift $z$
for each galaxy to $ugriz$ bands at some fixed redshift $z_0$. While
traditionally $K$-corrections are performed to $z_0=0$, this is not
always desirable, particularly for surveys like the SDSS for which
most of the galaxies are not observed near $z=0$.

Given AB galaxy magnitudes $\band{z}{m}$ and errors for a galaxy at
redshift $z$, one can use the method of the previous subsection to
reconstruct the galaxy rest-frame SED.  With this reconstruction in
hand, it is straightforward to project the SED onto any bands at any
redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
\begin{eqnarray}
\band{z_0}{m_{\mathrm{rec}}} &=& -2.41 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)} {\int_{0}^{\infty}
d\lambda \lambda^{-1} R(\lambda)}\right] \cr &=& -48.60 - 2.5
\log_{10}\left[ \frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)} {\int_{0}^{\infty} d\nu \nu^{-1}
R(\nu)}\right]
\end{eqnarray}
The simple way to remember where the factors of $(1+z)$ appear in
these equations is to remember that $K$-corrections are defined to be
the change in magnitude with redshift for {\it fixed bolometric flux}.
So when the spectrum is stretched by the $(1+z_0)$ redshift factor, as
in $f(\lambda/(1+z_0))$, the bolometric flux should be constant; that
is, you must apply the extra multiplicative factor $1/(1+z_0)$ to
counteract the stretching of the spectrum.

Given the reconstructed magnitudes, one can calculate the fixed frame
magnitudes in one of two ways:
\begin{enumerate}
\item Simply use $\band{z_0}{m_{\mathrm{rec}}}$, the reconstructed
magnitudes at redshift $z_0$, as the fixed frame magnitudes.
\item Use the difference
$\band{z_0}{m_{\mathrm{rec}}}-\band{z}{m_{\mathrm{rec}}}$ as an
estimate of the correction to apply to the observed $\band{z}{m}$.
\end{enumerate}
If the reconstructions $\band{z}{m_{\mathrm{rec}}}$ are very similar
to the actual observations $\band{z}{m}$ (as we will show is the case
for our fits), there is very little difference between these two
methods. However, if the reconstructions are poor, there will be large
differences between the methods; in particular, method (1) will yield
a color distribution of galaxies with artificially low
dimensionality. Having estimated $\band{z_0}{m}$, you can then
calculate the luminosity by simply applying to the calculated
magnitude the cosmological distance modulus (that is, the luminosity
distance-squared law, as tabulated by, {\it e.g.},  \citealt{hogg99a}).

@@

MB and DWH acknowledge NASA NAG5-11669 for partial support.  MB was
supported at the beginning of this work by the DOE and NASA grant NAG
5-7092 at Fermilab. He is also grateful for the hospitality of the
Department of Physics and Astronomy at the State University of New
York at Stony Brook, who kindly provided computing facilities on his
frequent visits there.

>>>>

\newcommand{\vv}[1]{{\bf #1}}
\newcommand{\df}{\delta}
\newcommand{\dfft}{{\tilde{\delta}}}
\newcommand{\betaft}{{\tilde{\beta}}}
\newcommand{\erf}{{\mathrm{erf}}}
\newcommand{\erfc}{{\mathrm{erfc}}}
\newcommand{\Step}{{\mathrm{Step}}}
\newcommand{\ee}[1]{\times 10^{#1}}
\newcommand{\avg}[1]{{\langle{#1}\rangle}}
\newcommand{\Avg}[1]{{\left\langle{#1}\right\rangle}}
\def\simless{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'074$}}\mathchar"7218\,$}}} % < or of order
\def\simgreat{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'076$}}\mathchar"7218\,$}}} % > or of order
\newcommand{\iras}{{\sl IRAS\/}}
\newcommand{\petroratio}{{{\mathcal{R}}_P}}
\newcommand{\petroradius}{{{r}_P}}
\newcommand{\petronumber}{{{N}_P}}
\newcommand{\petroratiolim}{{{\mathcal{R}}_{P,\mathrm{lim}}}}
\newcommand{\band}[2]{\ensuremath{^{{#1}}\!{#2}}}

@@

In the past, people have accounted for these effects using
``$K$-corrections'' applied to the observed magnitudes. In these
analyses, a function $K(z)$ is added to the standard cosmological
bolometric distance modulus $\mathrm{DM}(z)$ to obtain the
relationship between the apparent magnitude $m$ and the absolute
magnitude $M$: 
\begin{equation}
\label{kcorrecteqn}
m = M + \mathrm{DM}(z) + K(z)
\end{equation}
Sometimes a single function has been applied regardless of the galaxy
type, though more recently it has become standard to use a discrete
set of $K(z)$ functions depending on galaxy type (based either on
morphology or spectral features). As described in \cite{oke68a} and
later papers, the function is based on the projection of an assumed
galaxy SED redshifted to $z$ onto the measured instrument response as
a function of wavelength (including the effects of the atmosphere, the
reflectivity of the mirrors, the filter transmission, and the response
of the CCD device or photographic emulsion). For example, for the
Sloan Digital Sky Survey (SDSS) filter system, \citet{frei94a} and
\citet{fukugita95a} have both presented results for the
$K$-corrections of galaxies. Each of these papers tabulates $K(z)$
(called $k(z)$ by \citealt{frei94a}) for galaxies of various
morphological types, based on galaxy spectrophotometry.\footnote{Note
that the description on p. 57 of \citet{binney98a} of the meaning of
$k(z)$ in \citet{frei94a} is incorrect; in fact, their $k(z)$ is
exactly equivalent to our $K(z)$.}

@@

To characterize $K$-corrections, consider again Equation
(\ref{kcorrecteqn}), which we here generalize to the case of relating
an apparent magnitude observed in band $R$ to an absolute magnitude
observed in band $Q$:
\begin{equation}
m_R = M_Q + \mathrm{DM}(z) + K_{QR}(z).
\end{equation}
The definition of an apparent AB magnitude is given by Equation
(\ref{ABdef}). An absolute AB magnitude is defined to be the apparent
AB magnitude which would be measured for a galaxy 10 pc distant, at
rest.  In the restricted case of AB magnitudes, the resulting
$K$-correction is given by:
\begin{equation}
K_{QR}(z) = -2.5 \log_{10} \left[
\frac{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
}{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) Q(\nu)
}
\right]
\end{equation}
A more general expression holds for other magnitude systems (such as
the Vega-relative magnitude system) and for the case of
$K$-corrections when $Q$ and $R$ are in different magnitude systems.

In practice, we will be concerned here primarily with transforming the
apparent magnitudes in band $R$ to a band $Q$ which is just the $R$
band shifted by a factor $(1+z_0)$, which we will denote
\begin{equation}
\band{z_0}{R}(\nu) \equiv R(\nu/(1+z_0))  
\end{equation}
read, ``$R$-band shifted to $z_0$.''

In the case of transforming $R$ to $\band{z_0}{R}$
\begin{equation}
K_{\band{z_0}{R}R}(z) = -2.5 \log_{10} \left[
\frac{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu) R(\nu)
}{
\int_0^{\infty} \frac{d\nu}{\nu} f(\nu/(1+z))/(1+z) R(\nu/(1+z_0))
}
\right]
\end{equation}
In the special case that $z=z_0$, one is transforming the magnitude of
an observed galaxy at $z_0$ to its rest-frame magnitude in a band
shifted by $z_0$. For that case
\begin{equation}
K_{\band{z_0}{R}R}(z_0) = -2.5 \log_{10} (1+z_0)
\end{equation}
independent of $f(\nu)$ (which is natural, since both magnitudes
sample the same region of $f(\nu)$). The fact that one's uncertainties
in $f(\nu)$ do not matter makes the use of the fixed frame magnitudes
shifted to the redshift of the object in question useful, when it is
possible.

%Given AB galaxy magnitudes $\band{z}{m}$ and errors for a galaxy at
%redshift $z$, one can use the method of the previous subsection to
%reconstruct the galaxy rest-frame SED.  With this reconstruction in
%hand, it is straightforward to project the SED onto any bands at any
%redshift $z_0$ ({\it cf.} Equation \ref{ABdef}):
%\begin{eqnarray}
%\band{z_0}{m_{\mathrm{rec}}} &=& -2.41 - 2.5 \log_{10}\left[
%\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z_0}
%f_{\mathrm{rec}}(\frac{\lambda}{1+z_0}) R(\lambda)} {\int_{0}^{\infty}
%d\lambda \lambda^{-1} R(\lambda)}\right] \cr &=& -48.60 - 2.5
%\log_{10}\left[ \frac{\int_{0}^{\infty} d\nu \nu^{-1} (1+z_0)
%f_{\mathrm{rec}}[\nu (1+z_0)] R(\nu)} {\int_{0}^{\infty} d\nu \nu^{-1}
%R(\nu)}\right]
%\end{eqnarray}
%The simple way to remember where the factors of $(1+z)$ appear in
%these equations is to remember that $K$-corrections are defined to be
%the change in magnitude with redshift for {\it fixed bolometric flux}.
%So when the spectrum is stretched by the $(1+z_0)$ redshift factor, as
%in $f(\lambda/(1+z_0))$, the bolometric flux should be constant; that
%is, you must apply the extra multiplicative factor $1/(1+z_0)$ to
%counteract the stretching of the spectrum.

%We here define a notation which we will use in subsequent papers for
%the magnitude in a band $b$ which we have shifted blueward by a factor
%$(1+z)$, namely:
%\begin{equation}
%\band{z}{b}
%\end{equation}
%read, ``$b$-band shifted to $z$.''  The goal of ``$K$-correction'' is
%to convert the magnitudes of a galaxy at redshift $z$ observed in the
%$ugriz$ bands at redshift $z$ to the magnitudes which would be
%measured if the galaxy were at redshift at $z=0$ and observed through
%the $ugriz$ bands shifted to some fixed redshift $z_0$. While
%traditionally $K$-corrections are performed to $z_0=0$ (that is, the
%bands are not shifted), this is not always desirable, particularly for
%surveys like the SDSS for which most of the galaxies are not observed
%near $z=0$.

In practice, one can calculate the fixed frame magnitudes in one of
two ways:
\begin{enumerate}
\item Simply reconstruct $\band{z_0}{R}$ from the model
$f_{\mathrm{rec}}(\nu)$ 
\item Calculate the $K$-correction above and apply it to the observed
$R$ magnitude
\end{enumerate}
If the reconstructions of the observed magnitudes $R$ are very similar
to the actual observations (as we will show is the case for our fits),
there is very little difference between these two methods. However, if
the reconstructions are poor, there will be large differences between
the methods; in particular, method (1) will yield a color distribution
of galaxies with artificially low dimensionality. Having estimated
$\band{z_0}{m}$, you can then calculate the luminosity by simply
applying to the calculated magnitude the cosmological distance modulus
(that is, the luminosity distance-squared law, as tabulated by, {\it
e.g.}, \citealt{hogg99a}).

@@

Special thanks to Ivan Baldry for pointing out an error in this
manuscript and in the distributed code.  MB and DWH acknowledge NASA
NAG5-11669 for partial support.  MB was supported at the beginning of
this work by the DOE and NASA grant NAG 5-7092 at Fermilab. He is also
grateful for the hospitality of the Department of Physics and
Astronomy at the State University of New York at Stony Brook, who
kindly provided computing facilities on his frequent visits there.
